Basics of Hardware Cache Optimization

### 2017-06-06 tex/methodology.tex
Draft section on locality, tweak stuff

\begin{comment}
Plenty of things can go wrong when \alts{profiling, measuring \alts{execution, run}
times}.  These are the ones I was aware of and tried to account for while doing
measurements for this \article.
\begin{itemize}
   \item \emph{CPU frequency scaling and boosting}.
      CPU frequencies are usually dynamic these days and automatically adjusted based on
      at least workload and temperature.  This should be disabled when measuring execution
      times.  I used the cpupower(1) program.
      % See <https://wiki.archlinux.org/index.php/CPU_frequency_scaling> and
      % <https://github.com/torvalds/linux/tree/master/tools/power/cpupower>.
      \begin{minted}[autogobble]{text}
         # cpupower frequency-set -g performance
         $ ./benchmark
         # cpupower frequency-set -g ondemand
      \end{minted}
      %stopzone
      % See [5].  Not a problem when measuring cycles.
   \item \emph{Interrupts and context switches.}  The process being timed has to share
      resources with other processes and the operating system.  A context switch will not
      only increase wall time, it will also increase cycles because of TLB flushes and
      cache evictions. % See <https://en.wikipedia.org/wiki/Context_switch>.

@@

\begin{minted}[autogobble]{text}
   $ gcc -march=native -O2 program.c
   # cpupower frequency-set -g ondemand
   # chrt -f 99 ocount -e CPU_CLK_UNHALTED ./a.out
\end{minted}
%stopzone

>>>>

\begin{comment}
Plenty of things can go wrong when \alts{profiling, measuring \alts{execution, run}
times}.  These are the ones I was aware of and tried to account for while doing
measurements for this \article.
\begin{itemize}
   \item \emph{CPU frequency scaling and boosting}.
      CPU frequencies are usually dynamic these days and automatically adjusted based on
      at least workload and temperature.  This should be disabled when measuring execution
      times.  I used the cpupower(1) program.
      % See <https://wiki.archlinux.org/index.php/CPU_frequency_scaling> and
      % <https://github.com/torvalds/linux/tree/master/tools/power/cpupower>.
      \begin{minted}[autogobble]{text}
         # cpupower frequency-set -g performance
         $ ./benchmark
         # cpupower frequency-set -g schedutil
      \end{minted}
      %stopzone
      % See [5].  Not a problem when measuring cycles.
   \item \emph{Interrupts and context switches.}  The process being timed has to share
      resources with other processes and the operating system.  A context switch will not
      only increase wall time, it will also increase cycles because of TLB flushes and
      cache evictions. % See <https://en.wikipedia.org/wiki/Context_switch>.

@@

\begin{minted}[autogobble]{text}
   $ gcc -march=native -O2 program.c
   # cpupower frequency-set -g schedutil
   # chrt -f 99 ocount -e CPU_CLK_UNHALTED ./a.out
\end{minted}
%stopzone
